Supplement to: Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-18 nips_2018.tex
updated super figure
Almost midnight commit

\parbox{0.58\textwidth}{
We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
domain-specific primitives that the algorithm itself discovers,
together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration
of top-down symbolic representations and bottom-up neural networks --
both of them learned -- could help make program induction systems more
generally useful for AI. Many directions remain open. Two immediate
goals 
are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in
some domains over the generic methods used here.
Another direction is to
explore DSL meta-learning: Can we find a \emph{single} universal
primitive set that could effectively bootstrap DSLs for new domains,
including the three domains considered,  but also many others?
}\hfill\begin{minipage}{0.4\textwidth}\centering
  \includegraphics[width=0.8\textwidth]{figures/geomCompiled.eps} 
  \captionof{figure}{
  Top left: hand crafted traning targets. Top right: examples of
  discovered compiled new programs. Bottom: compiled programm across iterations
  to highlight structure emergence.
  }\label{geomCompiled}
\end{minipage}

>>>>

\parbox{0.58\textwidth}{
We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
domain-specific primitives that the algorithm itself discovers,
together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration
of top-down symbolic representations and bottom-up neural networks --
both of them learned -- could help make program induction systems more
generally useful for AI. Many directions remain open. Two immediate
goals 
are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in
some domains over the generic methods used here.
Another direction is to
explore DSL meta-learning: Can we find a \emph{single} universal
primitive set that could effectively bootstrap DSLs for new domains,
including the three domains considered,  but also many others?
}\hfill
\begin{minipage}{0.4\textwidth}
  \centering
  \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
  \captionof{figure}{
  Top left: a few example of the shapes used as targets. Top right: a few
  examples of discovered new shapes. Bottom: learning curve across iterations in
  percent in the middle, and on both side average of the compiled new shapes
  before any traning and at the end of the last iteration. See supplementary
  material for more details.}
\label{geomCompiled}
\end{minipage}

### 2018-05-18 supplement.tex
Almost midnight commit

\section{Learning abstract geometry}
The brifly presented geometry project is a recent addition to the set of
problems adressed with our method. The original DSL, while inspired by the
Logo~\cite{abelson1974logo} in the sense that ones controls a moving on-off
point, was designed to be congitively plausible in the sense that some
abstractions are baked in for curves and a few structures --- it was first
described in~\cite{sablemeyer2017geom}.

@@

\begin{center}
\tt
\begin{tabular}{rr}
  \toprule
  \textrm{\emph{name}}    & \textrm{\emph{type}} \\
  \midrule
  Concat    & \text{prog}      \to \text{prog}      \to \text{prog} \\
  Repeat    & var option        \to \text{prog}      \to \text{prog} \\
  Embed     & \text{prog}      \to \text{prog}                       \\
  Define    & \text{var}        \to \text{prog}                       \\
  Turn      & \text{var option} \to \text{prog}                       \\
  Integrate & \text{var option} \to \text{bool}       \to              \\
            & \text{var option} \to \text{var option} \to              \\
            & \text{prog}                                             \\
  \midrule
  True      & bool                                                     \\
  False     & bool                                                     \\
  \midrule
  Nothing   & \text{var option}                                        \\
  Just      & \text{var}        \to \text{var option}                  \\
  \midrule
  Unit      & var                                                      \\
  Name      & var                                                      \\
  Next      & \text{var}        \to \text{var}                         \\
  Prev      & \text{var}        \to \text{var}                         \\
  Double    & \text{var}        \to \text{var}                         \\
  Half      & \text{var}        \to \text{var}                         \\
  Opposite  & \text{var}        \to \text{var}                         \\
  \bottomrule
\end{tabular}
\end{center}

In this context programs to be learned are directly resulting shapes, since they
don't take an input, and all tasks will be unique in the sense that each program
ignores input and produces a single output.

Some elements of the semantics are common, the others are as follow.
\texttt{Repeat} takes a \texttt{variable} and a \texttt{prog} and repeats
said \texttt{prog} $n$ times where $n$ is the evaluation of the
\texttt{variable} --- if not set it is defaulted to two. \texttt{Embed} of a
\texttt{prog} means that said \texttt{prog} will be executed and then returns to
the current state --- leaving what has been drawn in the meantime on the canvas.
\texttt{Integrate} is the main instruction and the only one that draws anything:
\texttt{Integrate(t,p,a,c)} takes a time \texttt{var}, a pen \texttt{bool}, an
acceleration \texttt{var} and an angular speed \texttt{var}, and it moves the
turtle according to these parameters --- with or without actually drawing
depending on the pen variable \texttt{p}.

\begin{wrapfigure}{r}{0.25\textwidth}
  \begin{center}
    \includegraphics[width = 0.25\textwidth]{figures/DSLGeom.eps}
  \end{center}\label{dslgeom}\caption{Some defaults for \texttt{Integrate}}
\end{wrapfigure}

The default values of these parameters are set such that \texttt{Integrate(nothing,
true, nothing, nothing)} draws a unit segment, \texttt{Integrate(nothing, true,
nothing, Just(Unit))} draws a circle of unit length, and
\texttt{Integrate(nothing, true, Just(Unit), Just(unit))} draws the first spire
of a spiral.

The training was done on the following set of tasks

\begin{figure}[h]
\centering
\includegraphics[width = \textwidth]{figures/geomAllTasks.png}
  \caption{The set of tasks for the geometry domain}
\end{figure}

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}
  \begin{scope}[shift={(2.5,0)}]
    \node[obs] at (3,3) (dx){$\mathcal{D}$};
    \node[latent] at (3.5,1.75) (zp){$p$};
    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
    \node[obs] at (3.5,0.7) (xp) {$x$};
    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Search: Infer $p$};
    \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
    \draw [->,red] (tx) -- (zp);
    \draw [->] (dx) -- (zp);
    \node at (nn) {
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
  \end{scope}

  \begin{scope}[shift={(0.7,-4.2)}]
    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
    \draw [->] (p2.south) -- (x2.north);
    \draw [->] (dt.south) -- (p2.north);

    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
    \draw [->] (p1.south) -- (x1.north);
    \draw [->] (dt.south) -- (p1.north);

    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
    \draw [->] (p3.south) -- (x3.north);
    \draw [->] (dt.south) -- (p3.north);

    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};

    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
    \draw [->] (zp.south) -- (xp.north);
    \plate {}{(zp)(xp)}{$x\in X$};

    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
  \end{scope}

  \begin{scope}[shift={(7.7,-4)}]
    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
    \node[latent] at (1.5,3) (t){$\theta$};
    \node[obs] at (1,1.75) (z){$p$};
    % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
    \node[obs] at (1,0.5) (x) {$x$};
    \edge {z}{x};
    \edge {d,t}{z};
    \plate {}{(z)(x)}{$x\in X$};
    \node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
  \end{scope}

\end{tikzpicture}
\end{figure}

\section{Generative model over the programs}
\begin{wrapfigure}{L}{0.5\textwidth}
  \begin{minipage}{0.49\textwidth}    
\begin{algorithm}[H]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
%%      \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
%%   \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
%%   \STATE \textbf{Output:} a program whose type unifies with $\tau$
%%   \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
%% \STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
      \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ \\
      \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Unify $\tau$ with yield$(\tau')$.
   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
 \STATE\textbf{where:}
 \STATE yield$(\tau) = \begin{cases}
   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
   \tau   &\text{ otherwise.}
 \end{cases}$ 
 \STATE  args$(\tau) = \begin{cases}
   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
   []   &\text{ otherwise.}
 \end{cases}$
\end{algorithmic}
\end{algorithm}
    \end{minipage}
  \end{wrapfigure}Alg.~\ref{programGenerativeModel} is a procedure for drawing
samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.

>>>>

\section{Learning abstract geometry}
The brifly presented geometry project is a recent addition to the set of
problems adressed with our method.

The original DSL, while inspired by the Logo~\cite{abelson1974logo} in the sense
that ones controls a moving on-off point, was designed to be congitively
plausible in the sense that some abstractions are baked in for curves and a few
structures --- it was first described in~\cite{sablemeyer2017geom}.

@@

\begin{center}
\tt
\begin{tabular}{rl}
  \toprule
  \textrm{\emph{name}}    & \textrm{\emph{type}} \\
  \midrule
  Concat    & \text{prog}      \to \text{prog}        \to \text{prog} \\
  Repeat    & var option        \to \text{prog}       \to \text{prog} \\
  Embed     & \text{prog}      \to \text{prog}                        \\
  Define    & \text{var}        \to \text{prog}                       \\
  Turn      & \text{var option} \to \text{prog}                       \\
  Integrate & \text{var option} \to \text{bool}       \to             \\
            & \text{var option} \to \text{var option} \to             \\
            & \text{prog}                                             \\
  \midrule
  True      & bool                                                    \\
  False     & bool                                                    \\
  \midrule
  Nothing   & \text{var option}                                       \\
  Just      & \text{var}        \to \text{var option}                 \\
  \midrule
  Unit      & var                                                     \\
  Name      & var                                                     \\
  Next      & \text{var}        \to \text{var}                        \\
  Prev      & \text{var}        \to \text{var}                        \\
  Double    & \text{var}        \to \text{var}                        \\
  Half      & \text{var}        \to \text{var}                        \\
  Opposite  & \text{var}        \to \text{var}                        \\
  \bottomrule
\end{tabular}
\end{center}

In this context programs to be learned correspond to a single shape since they
don't take an input. All tasks will ignore input and produces a single output.
This differs from the previous examples where each target could be evaluated on
various input.

Some elements of the semantics are common, the others are as follow.
\texttt{Repeat} takes a \texttt{variable} and a \texttt{prog} and repeats
said \texttt{prog} $n$ times where $n$ is the evaluation of the
\texttt{variable} --- if not set it is defaulted to two. \texttt{Embed} of a
\texttt{prog} means that said \texttt{prog} will be executed and then returns to
the current state --- leaving what has been drawn in the meantime on the canvas.

\begin{wrapfigure}{r}{0.25\textwidth}
  \begin{center}
    \includegraphics[width = 0.25\textwidth]{figures/DSLGeom.eps}
  \end{center}\label{dslgeom}\caption{Top: Some defaults for \texttt{Integrate}.
  Bottom, the result of the \texttt{Cross} example bellow.}
\end{wrapfigure}

\texttt{Integrate} is the main instruction and the only one that draws anything:
\texttt{Integrate(t,p,a,c)} takes a time \texttt{var}, a pen \texttt{bool}, an
acceleration \texttt{var} and an angular speed \texttt{var}, and it moves the
turtle according to these parameters --- with or without actually drawing
depending on the pen variable \texttt{p}.

The default values of these parameters are set such that \texttt{Integrate(nothing,
true, nothing, nothing)} draws a unit segment, \texttt{Integrate(nothing, true,
nothing, Just(Unit))} draws a circle of unit length, and
\texttt{Integrate(nothing, true, Just(Unit), Just(unit))} draws the first spire
of a spiral. Playing with the first arguments decides on the duration during
which the arguments are integrated.

In the \texttt{var} type, most are self explicit, while \texttt{Opposite(v)}
takes the opposite of \texttt{v}. The behaviour of \texttt{Name} was originally
designed to handle arbitrary variables in a call by name fashion but was latter
reduced to a single storage location, which proves to be enough for the
targeted shapes. One can therefore store elements in that placeholder using
\texttt{Define} and retrieve it through \texttt{Name}.

Because of these defaults, the following program draws a cross:

\[
  \text{\texttt{Cross = Repeat(Double(Double(Unit)), Concat(Embed(Integrate), Turn(None)))}}
\]

Describing highly regular complex shapes in this language is easy to do as a
human but quickly escapes the reach of naïve enumeration search. By using a
curriculum of shapes our approach compresses the search in the corresponding
directions --- another way to say this is that upon being given a dataset of
shapes, it picks up the simple ones and abstract them as building blocks for
latter staged of search.

For example, the simplest way to draw a square in this language is already of
length 12. Placing several around, for example to draw a grid is out of reach of
the initial search. However by first abstracting as a primitive the segment ---
thus reducing the length by four ---, then assuming that after a segment it
often needs to draw something else, then abstracting the first arguments of
\texttt{Repeat} --- further reducing the length by two --- as well as the one of
\texttt{Turn} and finally making the square a primitive on its own once it
starts using it often enough, the length of this particular shape drops.

\begin{table}[htbp]
  \begin{center}
    \begin{tabular}{rcl}
      \toprule
      New Primitive & Type & Definition \\
      \midrule
      Segment = $f_0$ & \texttt{prog} & \texttt{Concat(Nothing, True, Nothing, Nothing)} \\
      Right-angle = $f_1$ & \texttt{prog} & \texttt{Turn(Nothing)} \\
      RepeatTwice = $f_2$ & \texttt{prog} $\to$ \texttt{prog} & $\lambda p.$ \texttt{Repeat(nothing,} $p$) \\
      AddToUnitSegment = $f_3$ & \texttt{prog} $\to$ \texttt{prog} & $\lambda p.$ \texttt{Concat(}$f_0, p$) \\
      Square = $f_4$ & \texttt{prog} & $f_3(f_3(f_2(f_1)))$ \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{How our method compresses the square step by step. On the example
  given in the main article this was produced by the compressor after the second
  search phase and leads to the second jump in success, the first one being the
  abstraction of the Segment. Names are \emph{not} produced by the compressor
  and are here as indication to help the reader.}
\end{table}

The underlying hypothesis is that the new primitive distort the space of search
toward something that looks more like what human actually produce and moves away
from semantically valid but meaningless programs --- in a sense, learns to care
about what \emph{matters} rather than what is \emph{true} in a very
pragmatic-like way.

In Figure \ref{alltasksgeom} are listed all the shapes used for this project
without particular order.

\begin{figure}[h]
\centering
\includegraphics[width = \textwidth]{figures/geomAllTasks.png}
\caption{The set of tasks for the geometry domain}
\label{alltasksgeom}
\end{figure}

Since this is a first step in broader project of program induction for abstract
geometry the likelyhood is currently all-or-none --- ongoing work moves this to
a neural net based distance function to abstract away from noise in the shapes.

\newpage

@@

\begin{figure}[h]
\centering
  \includegraphics{figures/iterations.eps}
%\begin{tikzpicture}
  %\begin{scope}[shift={(2.5,0)}]
    %\node[obs] at (3,3) (dx){$\mathcal{D}$};
    %\node[latent] at (3.5,1.75) (zp){$p$};
    %\node[obs] at (4,3) (tx){$\theta^{(x)}$};
    %\node[obs] at (3.5,0.7) (xp) {$x$};
    %\node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Search: Infer $p$};
    %\draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
    %\draw [->,red] (tx) -- (zp);
    %\draw [->] (dx) -- (zp);
    %\node at (nn) {
      %\begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        %\tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        %\fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        %\node[rectangle] at (1,1) {};
        %\foreach \name / \y in {1,...,4}
            %\node[neuron] (I-\name) at (0,-\y) {};
        %\foreach \name / \y in {1,...,3}
            %\node[neuron] (H-\name) at (1,-\y-0.5) {};
        %\foreach \name / \y in {1,...,4}
            %\node[neuron] (O-\name) at (2,-\y) {};
        %\foreach \source in {1,...,4}
            %\foreach \dest in {1,...,3}
                %\draw [-latex] (I-\source) -- (H-\dest);
        %\foreach \source in {1,...,3}
            %\foreach \dest in {1,...,4}
                %\draw [-latex] (H-\source) -- (O-\dest);
      %\end{tikzpicture}
    %};
    %\node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
  %\end{scope}

  %\begin{scope}[shift={(0.7,-4.2)}]
    %\node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
    %\draw [->] (p2.south) -- (x2.north);
    %\draw [->] (dt.south) -- (p2.north);

    %\node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
    %\draw [->] (p1.south) -- (x1.north);
    %\draw [->] (dt.south) -- (p1.north);

    %\node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
    %\draw [->] (p3.south) -- (x3.north);
    %\draw [->] (dt.south) -- (p3.north);

    %\node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
    %\node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};

    %\node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
    %\node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
    %\draw [->] (zp.south) -- (xp.north);
    %\plate {}{(zp)(xp)}{$x\in X$};

    %\draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
    %rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
    %\node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
  %\end{scope}

  %\begin{scope}[shift={(7.7,-4)}]
    %\node[latent] at (0.5,3) (d){$\mathcal{D}$};
    %\node[latent] at (1.5,3) (t){$\theta$};
    %\node[obs] at (1,1.75) (z){$p$};
     %\node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
    %\node[obs] at (1,0.5) (x) {$x$};
    %\edge {z}{x};
    %\edge {d,t}{z};
    %\plate {}{(z)(x)}{$x\in X$};
    %\node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
  %\end{scope}

%\end{tikzpicture}
\end{figure}

\section{Generative model over the programs}

Alg.~\ref{programGenerativeModel} is a procedure for drawing
samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.

\begin{wrapfigure}{L}{0.5\textwidth}
  \begin{minipage}{0.49\textwidth}    
\begin{algorithm}[H]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
%%      \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
%%   \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
%%   \STATE \textbf{Output:} a program whose type unifies with $\tau$
%%   \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
%% \STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
      \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ \\
      \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Unify $\tau$ with yield$(\tau')$.
   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
 \STATE\textbf{where:}
 \STATE yield$(\tau) = \begin{cases}
   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
   \tau   &\text{ otherwise.}
 \end{cases}$ 
 \STATE  args$(\tau) = \begin{cases}
   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
   []   &\text{ otherwise.}
 \end{cases}$
\end{algorithmic}
\end{algorithm}
    \end{minipage}
  \end{wrapfigure}
